---
title: "Using grattantheme"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using grattantheme}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}

height <- 4
width <- 6

knitr::opts_chunk$set(collapse = T, comment = "#>", 
                      fig.align='center',
                      fig.height = height,
                      fig.retina = 2,
                      fig.width = width)
library(ggplot2)
library(grattantheme)

```


This vignette explains how to use `grattantheme` to quickly and consistently apply Grattan chart formatting to charts made in `R` using `ggplot`.

When creating a chart using `ggplot` we have to:

* Choose a dataset;
* Map variables to chart aesthetics `aes()`;
* Choose a `geom_`.

For example, using the in-built `iris` dataset:

```{r}
plot <- ggplot(iris,
               aes(x = Sepal.Length,
                   y = Sepal.Width,
                   colour = Species)) +
        geom_point(size = 4) +
        labs(x = "Species",
             y = "",
             colour = "Species")
```

This successfully plots the data we want to plot:

```{r, echo = FALSE}
plot
```

But it doesn't yet _look_ like a Grattan chart. To adjust the _look_ we adjust '`theme`' elements, like `axis.ticks.x = element_line(colour = "black")` to adjust the axis tickmarks on the x axis; `panel.grid.major.x = element_blank()` to turn off vertical gridlines; and so on; and on; and on. We also need to adjust aesthetic colours to the Grattan palette; setting, for example, `fill = "#F68B33"`. The `grattantheme` package contains tools and shortcuts to simplify this process.


\newpage 
## Formatting theme elements with `theme_grattan()`

The function `theme_grattan()` contains all of the Grattan `theme` adjustments in one handy command. Combined with `grattan_colour_manual`, which easily changes colours of aesthetics, your `R` chart will be ready for a report or a slide in no time.


```{r}
plot + 
  theme_grattan()
```

By default, `theme_grattan()` supresses the legend to allow for clearer on-chart labelling. We can include the legend with the `legend` argument, which takes `"off"`, `"top"`, `"bottom"`, `"left"` or `"right"`:

```{r}
plot + 
  theme_grattan(legend = "top")
```

To align the y-axis with zero, change the y scale with `grattan_y_continuous()`:

```{r}
plot + 
  theme_grattan() +
  grattan_y_continuous()
```


Sometimes we'll want a chart for a box in a report. We can change the background colour with the `background` argument:

```{r}
plot + 
  theme_grattan(background = "box") +
  grattan_y_continuous()
```


The standard Grattan rules for `x` and `y` axes flip if the chart is a horizontal bar chart. The `x` axis then follows the rules of the `y` axis, and vice-versa. If we are using a 'flipped' chart (imlemented with `coord_flipped()`), we can tell `theme_grattan` this is the case using the argument `fillped` set to `TRUE`.

```{r}
plot + 
  coord_flip() +
  theme_grattan(flipped = TRUE) +
  grattan_y_continuous()
```


The final adjustments we can specify with `theme_grattan` are the font size and font family. The defaults meet Grattan formatting requirements, but if we do need to change them we can:

```{r}
plot + 
  theme_grattan(base_size = 8, base_family = "serif") +
  grattan_y_continuous()
```


\newpage
## Using Grattan colours

Grattan's colours are loaded with `grattantheme`. The HEX codes for individual Grattan colours can be called using `grattan_[colourname]`, eg `grattan_lightorange`. Colours names are taken from the chart-guide and are:

```{r, out.width = "250px", echo = FALSE}
knitr::include_graphics("grattan_pal.pdf")
```

We can call single colours:

```{r}
plot + 
  geom_hline(yintercept = 3, colour = grattan_orange) +
  geom_vline(xintercept = 6, colour = grattan_darkorange) +
  theme_grattan() +
  grattan_y_continuous()
```


We can also use the `scale_fill_grattan()` or `scale_colour_grattan()` functions to change the colours of our fill or colour _aesthetics_. These can be used for discrete/catagorical data (the default) or continuous data. 

### Discrete colours
In our example, we have three different flowers each represented by a colour. So we need to set `discrete = TRUE`

```{r}
plot + 
  theme_grattan() +
  grattan_y_continuous() +
  scale_colour_grattan(discrete = TRUE)

```

Note that if you need more than 10 colours your chart will not render some of the data. You can set a manual theme using `make_grattan_pal`

```{r}
ggplot(mtcars, aes(x = mpg, y = hp, colour = as.character(wt))) +
  geom_point() +
  scale_colour_manual(values = make_grattan_pal()(29))
```

### Continuous colours

`scale_(colour|fill)_grattan` includes an option for continuous colours: `discrete = FALSE`. 

```{r}
plot2 <- ggplot(iris,
               aes(x = Sepal.Length,
                   y = Sepal.Width,
                   colour = Sepal.Length)) +
        geom_point(size = 5) +
        grattan_y_continuous() +
        labs(x = "")

plot_f <- plot2 +
  theme_grattan() +
  scale_colour_grattan(discrete = FALSE)

plot_f
```

## Saving plots with `grattan_save`


The `grattan_save()` function saves your chart in a variety of types. We specify the type with the `type` argument that can take the arguments:

* `"normal"`: a standard report chart size, and the default. [**height** = 14.5cm x **width** = 22.16cm]
* `"wholecolumn`: a taller whole-column chart for reports. [22.16 x 22.16]
* `"fullpage"`: a full-page chart for reports. [22.16 x 44.32]
* `"tiny"`: a rarely-used short chart for reports. [11.08 x 22.16]
* `"fullslide"` : to produce charts that look like a full slide with the Grattan logo. This option allows notes and sources to accompany the saved figure. 
* `"fullslide_169"` : as with "fullslide", but in 16:9 orientation.
* `"blog"` : Like "fullslide" plots, but with narrower borders -- use this type for the Grattan Blog or other online outlets like The Conversation.

You can specify `type = "all"` to save in all formats.

The argument `filename` is required. You must include an extension; `.pdf` is standard for Grattan charts for reports; `.png` is standard for the Grattan Blog. 

`grattan_save()` uses the ggplot2 function `ggsave()` to save your chart. Like `ggsave()`, `grattan_save()` will use the last plot you displayed by default, but you can specify something else with `object`.

Now we can save our Grattan-formatted graph as a normally-formatted report chart:

```{r}
# Create and store the plot:
plot_final <- plot_f +
              labs(title = "Iris plants are rad!",
                 subtitle = "Width of sepal",
                 x = "Length of sepal", 
                 y = "",
                 caption = "Notes: A classic dataset. Source: Fisher (1936)")

# Save the plot
grattan_save(filename = "iris.pdf",
             object = plot_final,
             type = "normal")

```


Which produces a chart that can fit into a report. Note that the title, subtitle and notes/sources have been removed.

```{r, out.width = "300px", echo = FALSE}
knitr::include_graphics("iris.pdf")
```

If we wanted to save a full-slide chart, complete with a title, subtitle, notes/sources and the Grattan slide format, we specify `type = "fullslide"`:

```{r, eval = FALSE}
# Save the plot
grattan_save(filename = "iris_fullslide.pdf",
             object = plot_final,
             type = "fullslide")

```

```{r, include=FALSE}
# Actually save the plot to call, but don't include output
grattan_save(filename = "iris_fullslide.pdf",
             object = plot_final,
             type = "fullslide")

```

```{r, out.width = "400px", echo = FALSE}
knitr::include_graphics("iris_fullslide.pdf")
```

### Saving Powerpoint files

It's convenient to also save your files in Powerpoint format. This allows PDs to easily use your graphs in their presentations. Saving in this format also means you can include speaker notes -- such as information about the script used to create the file -- which can help us to keep track of things.

You can save a Powerpoint file, with appropriate speaker notes included, like this:

```{r, eval = FALSE}
grattan_save_pptx(plot_final,
                  "iris_fullslide.pptx",
                  type = "fullslide")
```

Or, you can just use the `save_pptx` argument to `grattan_save()`, like this:

```{r, eval = FALSE}
grattan_save("iris_fullslide.pdf",
             plot_final,
             type = "fullslide",
             save_pptx = TRUE)
```

### Saving chart data

When we release reports or working papers, we typically upload an Excel file including 'chart data' for most of our charts. The process for making and properly formatting these files can be a bit laborious. `grattantheme` helps streamline that process, using the `save_chartdata()` function. 

You can use it like this:

```{r, eval = FALSE}
save_chartdata("plot_final.xlsx", object = plot_final)
```

Or, just use the `save_data` argument to `grattan_save()` when you save your plot:
```{r, eval = FALSE}
grattan_save("plot_final.png",
             type = "fullslide",
             save_data = TRUE)
```
