---
title: "Using grattantheme"
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Using grattantheme}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(ggplot2)
library(grattantheme)
```


```{r, eval = FALSE, include = FALSE}
# Run this to build the vignette:
devtools::build_vignettes()
```


When creating a chart in `R` you have to:

* Choose a dataset;
* Map variables to chart aesthetics `aes()`;
* Choose a `geom_`.

For example, using the `Orange` dataset tracking the growth of five orange trees by age:

```{r}
plot <- ggplot(Orange,
               aes(x = age,
                   y = circumference,
                   colour = Tree)) +
        geom_line() +
        labs(x = "Age of tree",
             y = "")
```

This successfully plots the data we want to plot:

```{r, echo = FALSE}
plot
```

But it doesn't _look_ like a Grattan chart. To adjust the _look_ we use `theme` elements, like `axis.ticks.x = element_line(colour = "black")` to adjust the axis tickmarks on the x axis, and `panel.grid.major.x = element_blank()` to turn off vertical gridlines, and so on, and on, and on. 

The function `theme_grattan()` contains all of the Grattan style-guide elements in one handy command. Combined with `grattan_colour_manual` to easily change colours of aesthetics, your `R` chart will be ready for a report or a slide in no time.

## Formatting theme elements with `theme_grattan()`

```{r}
plot + 
  theme_grattan()
```

By default, `theme_grattan()` supresses the legend to allow for clearer on-chart labelling. We can include the legend with the `legend` argument, which takes `"off"`, `"top"`, `"bottom"`, `"left"` or `"right"`:

```{r}
plot + 
  theme_grattan(legend = "top")
```



## Using Grattan colours

Grattan's colours are loaded with `grattantheme`. It has all the classics: \dots

We can use the `grattan_colour_manual` function to change the colours of our colour aesthetic. In our example, we have five different trees each represented by a colour, so we set the number of colours to five `grattan_colour_manual(n = 5)`:

```{r}
plot + 
  theme_grattan() +
  grattan_colour_manual(n = 5)
```

We can reverse the order of the colours using the `reverse` argument:

```{r}
plot + 
  theme_grattan() +
  grattan_colour_manual(n = 5, reverse = TRUE)
```




