---
title: "Using grattantheme"
output:
  pdf_document:
    toc: TRUE
    fig_width: 4
    fig_height: 3
vignette: >
  %\VignetteIndexEntry{Using grattantheme}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>", fig.align='center')
library(ggplot2)
library(grattantheme)
```


This vignette explains how to use `grattantheme` to quickly and consistently apply Grattan chart formatting to charts made in `R` using `ggplot`.

When creating a chart using `ggplot` we have to:

* Choose a dataset;
* Map variables to chart aesthetics `aes()`;
* Choose a `geom_`.

For example, using the `Orange` dataset tracking the growth of five orange trees by age:

```{r}
plot <- ggplot(Orange,
               aes(x = age,
                   y = circumference,
                   fill = Tree)) +
        geom_bar(stat = "identity") +
        labs(x = "Age of tree",
             y = "",
             colour = "Tree")
```

This successfully plots the data we want to plot:

```{r, echo = FALSE}
plot
```

But it doesn't yet _look_ like a Grattan chart. To adjust the _look_ we adjust '`theme`' elements, like `axis.ticks.x = element_line(colour = "black")` to adjust the axis tickmarks on the x axis; `panel.grid.major.x = element_blank()` to turn off vertical gridlines; and so on; and on; and on. We also need to adjust aesthetic colours to the Grattan palette; setting, for example, `fill = "#F68B33"`. The `grattantheme` package contains tools and shortcuts to simplify this process.


\newpage 
## Formatting theme elements with `theme_grattan()`

The function `theme_grattan()` contains all of the Grattan `theme` adjustments in one handy command. Combined with `grattan_colour_manual`, which easily changes colours of aesthetics, your `R` chart will be ready for a report or a slide in no time.


```{r}
plot + 
  theme_grattan()
```

By default, `theme_grattan()` supresses the legend to allow for clearer on-chart labelling. We can include the legend with the `legend` argument, which takes `"off"`, `"top"`, `"bottom"`, `"left"` or `"right"`:

```{r}
plot + 
  theme_grattan(legend = "top")
```

To align the y-axis with zero, change the y scale with `grattan_y_continuous()`:

```{r}
plot + 
  theme_grattan() +
  grattan_y_continuous()
```


Sometimes we'll want a chart for a box in a report. We can change the background colour with the `background` argument:

```{r}
plot + 
  theme_grattan(background = "box") +
  grattan_y_continuous()
```


The standard Grattan rules for `x` and `y` axes flip if the chart is a horizontal bar chart. The `x` axis then follows the rules of the `y` axis, and vice-versa. If we are using a 'flipped' chart (imlemented with `coord_flipped()`), we can tell `theme_grattan` this is the case using the argument `fillped` set to `TRUE`.

```{r}
plot + 
  coord_flip() +
  theme_grattan(flipped = TRUE) +
  grattan_y_continuous()
```


The final adjustments we can specify with `theme_grattan` are the font size and font family. The defaults meet Grattan formatting requirements, but if we do need to change them we can:

```{r}

plot + 
  theme_grattan(base_size = 8, base_family = "serif") +
  grattan_y_continuous()

```


\newpage
## Using Grattan colours

Grattan's colours are loaded with `grattantheme`. The HEX codes for individual Grattan colours can be called using `grattan_[colourname]`, eg `grattan_lightorange`. Colours names are taken from the chart-guide and are:

```{r, out.width = "250px", echo = FALSE}
knitr::include_graphics("grattan_pal.pdf")
```

We can call a single colour:

```{r}
plot + 
  geom_vline(xintercept = 850, colour = grattan_grey3) +
  theme_grattan() +
  grattan_y_continuous()

```


We can also use the `grattan_fill_manual()` or `grattan_colour_manual()` functions to change the colours of our fill or colour _aesthetics_. In our example, we have five different trees each represented by a colour, so we set the number of colours to five: `grattan_fill_manual(n = 5)`:

```{r}
plot + 
  theme_grattan() +
  grattan_y_continuous() +
  grattan_fill_manual(n = 5)

```

We can reverse the order of the fill colours using the `reverse` argument:

```{r}
plot + 
  theme_grattan() +
  grattan_y_continuous() +
  grattan_fill_manual(n = 5, reverse = TRUE)

```

Note that if you do not specify _enough_ colours, will receive an error:

```{r, eval = TRUE, error= TRUE, fig.show='hide'}
plot + 
  theme_grattan() +
  grattan_y_continuous() + 
  grattan_fill_manual(n = 3)
```

You won't receive an error if you specify _too many_ colours. `ggplot` will choose the first `n` colours it needs and ignore the rest.


## Saving plots with `grattan_save`


The `grattan_save()` function implements `ggsave()` with Grattan defaults for different occasions. We specify the occasion with the `type` argument that can take the arguments:

* `"normal"`: a standard report chart size, and the default. [**height** = 14.5cm x **width** = 22.16cm]
* `"wholecolumn`: a taller whole-column chart for reports. [22.16 x 22.16]
* `"fullpage"`: a full-page chart for reports. [22.16 x 44.32]
* `"tiny"`: a rarely-used short chart for reports. [11.08 x 22.16]
* `"fullslide"` : to produce charts on a full slide with the Grattan logo---to be used in articles or blogposts---use `"type = fullslide"`. This option allows notes and sources to accompany the saved figure. 

The argument `filename` is required with an extension (`.pdf` is standard for Grattan charts). Like `ggsave`, the `ggplot2` object to be saved will default to the last plot, but you can specify something else with `object`.

Now we can save our Grattan-formatted graph as a normally-formatted report chart:

```{r}
# Create and store the plot:
plot_final <- plot +
              theme_grattan() +
              grattan_y_continuous() +
              grattan_fill_manual(n = 5, reverse = TRUE) +
              labs(title = "Intergenerational inequality in trees",
                 subtitle = "Circumference of trees by age",
                 x = "Age", 
                 y = "",
                 caption = "Notes: orange trees only. Source: inbuilt base R datasets.")

# Save the plot
grattan_save(filename = "tree_plot.pdf",
             object = plot_final,
             type = "normal")

```


Which produces a chart that can fit into a report. Note that the title, subtitle and notes/sources have been removed.

```{r, out.width = "300px", echo = FALSE}
knitr::include_graphics("tree_plot.pdf")
```

If we wanted to save a full-slide chart, complete with a title, subtitle, notes/sources and the Grattan slide format, we specify `type = "fullslide"`:

```{r, eval = FALSE}
# Save the plot
grattan_save(filename = "tree_plot_fullslide.pdf",
             object = plot_final,
             type = "fullslide")

```

```{r, include=FALSE}
# Actually save the plot to call, but don't include output
grattan_save(filename = "tree_plot_fullslide.pdf",
             object = plot_final,
             type = "fullslide")

```

```{r, out.width = "400px", echo = FALSE}
knitr::include_graphics("tree_plot_fullslide.pdf")
```






